We want to acknowledge the use of a small number of routines,
written by other authors, in developing the siesta code.
In most cases, these routines were acquired by now-forgotten
routes, and the reported authorships are based on their headings.
If you detect any incorrect or incomplete attribution, or suspect
that other routines may be due to different authors, please
let us know.

\begin{itemize}
  \item%
  The main nonpublic contribution, that we thank thoroughly, are
  modified versions of a number of routines, originally written by
  \textbf{A. R.\ Williams} around 1985, for the solution of the radial
  Schr\"odinger and Poisson equations in the APW code of Soler and
  Williams (PRB \textbf{42}, 9728 (1990)).  Within \siesta, they are
  kept in files arw.f and periodic\_table.f, and they are used for the
  generation of the basis orbitals and the screened pseudopotentials.

  \item%
  The exchange-correlation routines contained in SiestaXC were written
  by J.M.Soler in 1996 and 1997, in collaboration with
  \textbf{C.\ Balb\'as} and \textbf{J. L.\ Martins}.  Routine pzxc,
  which implements the Perdew-Zunger LDA parametrization of xc, is
  based on routine velect, written by \textbf{S.\ Froyen}.

  \item%
  The serial version of the multivariate fast fourier transform used
  to solve Poisson's equation was written by \textbf{Clive Temperton}.

  \item%
  Subroutine iomd.f for writing MD history in files was originally
  written by \textbf{J. Kohanoff}.

\end{itemize}

We want to thank very specially \textbf{O. F.\ Sankey}, \textbf{D. J.\
    Niklewski} and \textbf{D. A.\ Drabold} for making the FIREBALL
code available to P.\ Ordej\'on.  Although we no longer use the
routines in that code, it was essential in the initial development of
the \siesta\ project, which still uses many of the algorithms
developed by them.

We thank \textbf{V. Heine} for his support and encouraging us in this
project.

The \siesta\ project is supported by the Spanish DGES through
several contracts. We also acknowledge past support by the Fundaci\'on
Ram\'on Areces.