The CheSS solver uses an expansion based on Chebyshev polynomials to calculate
the density matrix, thereby exploiting the sparsity of the overlap
and Hamiltonian matrices.
It works best for systems exhibiting a finite HOMO-LUMO gap and a small spectral width.

CheSS exhibits a two level parallelization using MPI and OpenMP and can scale
to many thousands of cores.
It can be downloaded and installed freely from
\url{https://launchpad.net/chess} (but note that this version might
not be compatible with the latest versions of the BigDFT lower-level libraries
on which it depends.

Note that this is still an experimental feature. Feedback is welcome.

\subsubsection{Input parameters}
Usually CheSS only requires little user input, as the default values for the input parameters
work in general quite well. Moreover CheSS has the capability to determine certain optimal values
on its own. The only input parameters which usually require some human action are the values
of the buffers required for the matrix multiplications to calculate the Chebyshev polynomials.

\begin{fdfentry}{CheSS!Buffer!Kernel}[length]<$4.0\,\mathrm{Bohr}$>
  Buffer for the density kernel within the CheSS calculation.
\end{fdfentry}

\begin{fdfentry}{CheSS!Buffer!Mult}[length]<$6.0\,\mathrm{Bohr}$>
  Buffer for the matrix vector multiplication within the CheSS calculation.
\end{fdfentry}

\begin{fdfentry}{CheSS!Fscale}[energy]<$10^{-1}\,\mathrm{Ry}$>
	Initial guess for the error function decay length (will be adjusted automatically).
\end{fdfentry}

\begin{fdfentry}{CheSS!FscaleLowerbound}[energy]<$10^{-2}\,\mathrm{Ry}$>
	Lower bound for the error function decay length.
\end{fdfentry}

\begin{fdfentry}{CheSS!FscaleUpperbound}[energy]<$10^{-1}\,\mathrm{Ry}$>
	Upper bound for the error function decay length.
\end{fdfentry}

\begin{fdfentry}{CheSS!evlowH}[energy]<$-2.0\,\mathrm{Ry}$>
	Initial guess for the lower bound of the eigenvalue spectrum of the Hamiltonian matrix, will be adjusted automatically if chosen unproperly.
\end{fdfentry}

\begin{fdfentry}{CheSS!evhighH}[energy]<$2.0\,\mathrm{Ry}$>
	Initial guess for the upper bound of the eigenvalue spectrum of the Hamiltonian matrix, will be adjusted automatically if chosen unproperly.
\end{fdfentry}

\begin{fdfentry}{CheSS!evlowS}[real]<$0.5$>
	Initial guess for the lower bound of the eigenvalue spectrum of the overlap matrix, will be adjusted automatically if chosen unproperly.
\end{fdfentry}

\begin{fdfentry}{CheSS!evhighS}[real]<$1.5$>
	Initial guess for the upper bound of the eigenvalue spectrum of the overlap matrix, will be adjusted automatically if chosen unproperly.
\end{fdfentry}
